---
output:
  govdown::govdown_document:
    font: "sans-serif" # (default) or "new-transport"
    favicon: "govuk" # or "none" (default) or "govuk" see ?govdown_document
    # logo: "images/logo.svg" # or false (default)
    logo_url: "https://ukgovdatascience.github.io/govdown/"
    logo_text: "GOV.UK"
    page_title: "Explore education statistics - SEN Key Statistics"
    title: "Explore education statistics"
    phase: beta
    feedback_url: "https://github.com/ukgovdatascience/govdown/issues"
    theme:
      bslib: true
---

<style>
shadowless{
    box-shadow:0px;
}
</style>


``` {r include = FALSE}
library(rmarkdown)
library(bslib)
library(govdown)
knitr::opts_chunk$set(echo = FALSE)
vbox_gov_theme <- value_box_theme(bg = "#f3f2f1")
```


## Special educational needs statistics

``` {r}
bslib::layout_columns(
  col_widths = c(6,6),
bslib::value_box(
"Number of pupils with SEN requriements", 
  "1,766,924", 
  theme = vbox_gov_theme,
  class = "shadowless"
  ),
bslib::value_box(
"Percentage of pupils with SEN requirements", 
  "19.5%", 
  theme = vbox_gov_theme
  )
)
```

Source: [Special educational needs in England, Academic year 2025](https://explore-education-statistics.service.gov.uk/find-statistics/special-educational-needs-in-england/2024-25)

## Overall SEN totals

``` {r}
bslib::layout_columns(
  col_widths = c(4, 4, 4),
  bslib::value_box(
    "Number of pupils on SEN support",
    "1,284,284",
    "Increase of 3.7%",
    theme = vbox_gov_theme
  ),
  bslib::value_box(
    "Number of Education, health and care plans at January 2025",
    "482,640",
    "Increase of 11.1%",
  theme = vbox_gov_theme
  ),
  bslib::value_box(
    "Total pupils with special education needs",
    "1,766,924",
    "Increase of 4.3%",
  theme = vbox_gov_theme
  )
)
```

Source: [Special educational needs in England, Academic year 2025](https://explore-education-statistics.service.gov.uk/find-statistics/special-educational-needs-in-england/2024-25)

### SEN classifications

#### SEN support

SEN support means support that is additional to, or different from, the 
support generally made for other children of the same age in schools. It is provided for pupils who are identified as having a learning difficulty or a disability that requires 
extra or different help to that normally provided as part of the school’s usual 
curriculum offer. A pupil on SEN Support will not have an Education, Health and Care 
plan.

#### Education, Health and Care (EHC) plan

A local authority may issue an EHC plan for a 
pupil who needs more support than is available through SEN Support. This will follow 
a statutory assessment process whereby the local authority considers the pupil’s 
special educational needs and any relevant health and social care needs; sets out 
long term outcomes; and specifies provision which will deliver additional support to 
meet those needs

## Type of need

``` {r}
bslib::layout_columns(
  col_widths = c(6,6),
  bslib::value_box(
    "Autistic spectrum disorder",
    "132,200 (33.0%)",
    "Most common primary need for pupils on EHC plans",
    theme = vbox_gov_theme
  ),
  bslib::value_box(
    "Speech, language and communication needs",
    "291,700 (26.6%)",
    "Most common primary need for pupils on SEN support",
    theme = vbox_gov_theme
  )
)
```

Source: [Special educational needs in England, Academic year 2025](https://explore-education-statistics.service.gov.uk/find-statistics/special-educational-needs-in-england/2024-25)

## Characteristics

### Gender

Special educational needs are more prevalent in boys than girls. 72.0% of pupils with an EHC plan in 2024 are boys, and 62.0% of pupils with on SEN support are boys. Both of 
these rates have been slowly decreasing in recent years.

### Free school meal eligibility

Pupils with special educational needs are more likely to be eligible for free school meals. 
42.2% of pupils with an EHC plan and 38.3% of pupils on SEN support were eligible for 
free school meals in January 2024 compared to 21.4% of pupils without identified special 
educational needs.

### Ethnicity

In January 2024, among the different ethnic groups, those recorded as 'Traveller of Irish 
heritage', 'Black Caribbean’ and ‘Other Black background’ had the highest proportion of 
pupils with an EHC plan, with percentages of 7.3%, 6.4% and 6.3% respectively.
Conversely, those recorded as ‘Chinese’ ethnic group had the lowest percentage of 
pupils with an EHC plan at 2.2%.

### English as a first language

84.2% of pupils who have SEN support and 84.2% of pupils with an EHC plan have a 
first language known to be or believed to be English. This is a slight decrease from the 
previous year and compares to 78.7% of all pupils in schools.

## Educational settings

### Schools

In January 2024, 81.3% of pupils with SEN were in state-funded mainstream school, 
similar to the January 2023 figure of 81.6%. Although the number of pupils with an EHC 
plan in special schools has increased between 2023 and 2024, the percentage of pupils 
with an EHC plan attending state-funded special schools has dropped from 37.9% to 
35.8%. These trends continue patterns seen since 2018. 

The percentage of pupils with an EHC plan attending independent schools has increased 
significantly in recent years. In January 2015, 5.3% of all pupils with statements attended 
independent schools, and this has increased to 7.7% of all pupils with an EHC plan in 
January 2024.

### Early years

In 2023, the number of children registered for the 15-hour entitlements was at the lowest 
point in each series, down to 1,196,0006. Of these children, the percentage all of 
children with SEN, that is both SEN support and EHCPs, increased from 7.0% in 2022 to 
8.0% in 2023.

### Looked after children and children in need with SEN – prevalence and 
characteristics

Prevalence: 58.1% of children who had been looked-after continuously for 12 months at 
31 March 2023 for whom data were available had SEN , which consists of 30.8% with an 
EHC plan and 27.3% with SEN support. This compares with 50.0% of children in need 
with SEN. In the overall pupil population, 17.1% of pupils are identified with SEN. 
The most common type of need for all looked after children at 31 March 2023 was 
‘Social, Emotional and Mental Health’

## Absence and exclusions

### Absence

Absence
12.3% of sessions were missed for pupils with EHC plans over 2022 to 2023 compared 
to 10.2% for pupils on SEN support and 6.6% for pupils without SEN. 
36.0% of pupils with EHC plans were persistent absentees8 over 2022 to 2023 compared 
to 31.1% for pupils on SEN support and 18.4% for pupils without SEN. 
2022 to 2023 was the second academic year which was entirely “post-pandemic” with no 
disruptions caused by periods of lockdown since 2018 to 2019, with the first being 2021 
to 2022. Due to the disruptions faced during previous years, caution should be take 
comparing current data with historical data.

### Exclusions

The rate of suspensions in the Spring term of 2022 to 2023 among pupils with EHC plans 
is 6.98, which is lower than for those on SEN support at 8.04. This compares to 2.18 for 
pupils with no SEN.
The rate of permanent exclusions in the same period among those pupils who have an 
EHC plan is 0.07, which, like suspensions, is lower than for those on SEN support at 
0.12. This compares to 0.02 for pupils with no SEN.

## Outcomes and attainment

### Early Years Foundation Stage Profile (EYFSP)

Pre COVID-19 pandemic, there was a steady increase over time in the percentage of 
children with SEN achieving a good level of development7 . Following the pandemic, in 
2022 to 2023, 19.8% of children with SEN were assessed at a good level of 
development, which is 1.0 percentage points higher than the figure for 2021 to 2022 of 
18.8%. Of these, 24.3% of children with SEN support had a good level of development, 
up from 22.9% and 3.8% of children with an EHCP did, up from 3.6%. The percentage of 
children with no identified SEN with a good level of development also increased to 74.0% 
in 2022 to 2023 from 70.9% in 2021 to 2022, an increase of 3.1 percentage points.

### Phonics screening check

The percentage of year 1 pupils who met the expected standard in the phonics screening 
check is similar to pre-pandemic levels. In 2022 to 2023, 42% of pupils with SEN met the 
expected standard in the latest measurement period, compared with 38% in 2021 to 
2022 and 43% in 2018 to 2019. Of those pupils with SEN, 48% of those on SEN support 
and 20% of those with an EHC plan met the expected standard in 2022 to 2023. The 
proportion of pupil with no identified SEN meeting the expected standard in the phonics

screening check is also similar to pre-pandemic levels, 86% in 2022 to 2023 compared 
to, 82% in 2021 to 2022 and 88% in 2018 to 2019. 

### Key Stage 1

The percentage of pupils who achieved the expected standard in KS1 teacher 
assessments (TA) has dropped in both SEN and non-SEN categories following the 
COVID-19 pandemic. Teacher assessments were not collected in 2019 to 2020 and 2020
to 2021 due to the COVID pandemic.

### Key Stage 2 

24% of pupils on SEN support and 8% of pupils with EHCPs achieved the expected level 
in reading, writing and mathematics at the end of KS2 in 2022 to 2023 compared to 70% 
of those with no identified SEN – these figures have increased compared to 2021 to 2022
where 21% of pupils on SEN support, 7% of pupils with an EHCP and 69% of pupils 
with no identified SEN reached the expected standard but are reductions in comparison 
to 2018 to 2019 when the figures were 25%, 9% and 74% respectively.
Pupils with SEN have lower progress scores compared to those with no identified SEN 
and the score is lowest in writing TA. 

### Key Stage 4 

The average attainment 8 score for pupils with SEN support in 2022 to 2023 was 33.3 
compared to 50.0 for those with no identified SEN. For those with an EHC plan this was 
14.0. For context, a higher attainment 8 score means greater attainment at the end of 
KS4.
In 2022 to 2023, 20.7% of pupils with SEN support achieved grades 5 or above in 
English and mathematics GCSEs, compared to 51.0% of pupils with no identified SEN. 
For those with an EHC plan this was 6.9%

### Post-16 – attainment by age 19

56.3% of pupils who were in state-funded mainstream schools and identified with SEN in 
year 11 achieved Level 2 (equivalent to 5+ 9-4 at GCSE) by age 19 in 2022 to 2023, 
which is 32.8 percentage points lower than pupils with no identified SEN (89.1%). 

## Post-16 learner participation

### Participation for 16- and 17-year-olds
Raising the participation age (RPA) legislation was introduced in 2013 to 2014 increasing 
the age to which all young people in England are required to remain in education or 
training. 87.2% of 16–17-year-olds flagged as having an EHC plan9 were in education 
and training in March 2023 compared with 93.2% of those flagged with having no 
identified SEN. These are similar to the previous year’s figures of 88.7% and 93.7% 
respectively.

### Participation in Further Education
In the Academic Year 2022 to 2023, out of all adults aged 19+ in government-funded 
further education and skills, learners recorded as having a Learning difficulty/disability or 
health problem (LLDD) account for 18.5% of the cohort (325,470) an increase of 11.8% 
from 291,010 in 2021 to 2022.

## Destinations

### After Key Stage 4 

Pupils with SEN at the end of KS4 were 5.7 percentage points less likely to have any 
sustained destination than those with no identified SEN. 88.9% of pupils with SEN went 
onto education, employment or apprenticeships compared to 94.6% of those with no 
identified SEN (2021 to 2022 destinations). To be counted in a destination, young people 
have to have sustained participation for a 6 month period in the destination year.

### After Key Stage 5 

Students with recorded special educational needs (SEN) in mainstream schools were 5.2 
percentage points less likely to have an overall sustained destination than those students 
without SEN (83.9% compared to 89.1%, respectively, for 2021 to 2022 destinations). 
69.8% of those with self-declared Learning Difficulty, Disability or health problem (LLDD) 
in state-funded mainstream colleges progressed to a sustained education, employment 
or apprenticeship destination compared to 78.5% of those without LLDD (2021 to 2022
destinations)

### Progression to higher education (HE)

In 2021 to 2022, 8.9% of pupils with an EHC plan at the end of KS4 progressed to HE by 
age 19, compared to 24.2% of pupils with SEN support at the end of KS4 and 51.2% for 
pupils with no identified SEN. The progression rate for pupils with an EHC plan, SEN 
support and pupils with no identified SEN all increased from 8.7%, 22.5% and 48.6% 
respectively in the previous year.

### Employment status 

In the 2022 to 23 financial year, 4.8% of adults with learning disabilities aged 18-64 who 
were receiving support from social services were in paid employment.

## Experience of the SEND system

### Total number of EHC plans maintained by local authorities

There were 576,000 children and young people with Education, Health and Care (EHC) 
plans as of January 202410. This is an increase of 11.4% from 517,000 as at January 
2023. This follows similar increases in recent years. 

### Number of new EHC plans issued by local authorities

84,400 new EHC plans were issued during 2023, an increase of 26.6% when compared 
against 2022. This is significantly higher than the increase of 11.0% from 2021 to 2022. 

### Number of initial requests for assessment

There were 138,200 initial requests made for assessment for an EHC plan during 2023, 
up from 114,500 in 2022 and the highest number since data was first collected in 2016. 

### Timeliness of issuing statements and EHC plans

Of those plans where a decision to issue a new plan was within the 2023 calendar year 
(excluding cases where exceptions apply), 50.3% were issued within the statutory 20 
week time limit. This is an increase from 2022 when 49.2% of new plans were issued 
within the time limit.
This figure excludes cases where under the legislation exceptional circumstances mean 
that the local authority need not comply with the 20 week time limit if it is impractical to do 
so.

### Appeals registered with the SEND tribunal

Parents/carers and young people can register an appeal with the SEND tribunal if they 
disagree with a decision their Local Authority (LA) has made regarding an EHC plan or 
needs assessment.There were 15,615 appeals registered in 2023 (calendar year), which
is equivalent to an estimated 2.5% of appealable decisions. This is an increase of 22% 
compared to the previous year when there were 12,796 appeals registered equivalent to 
an estimated 2.3% of appealable decisions

### Disability

The DWP Family Resources Survey shows that in 2022-23, 11% of children in the UK 
had a disability. This proportion has increased from 6% in 2011-12. For disabled children, 
the most common impairment type was social or behavioural impairments, with half 
reporting this. This was followed by a learning impairment, at 32%; then a mental health 
impairment, at 30% 
The DfE Childcare and Early Years survey of Parents for 2021 suggests that 8% of 
children have longstanding health conditions or disabilities. Children in lone parent 
families are more likely to have a long-standing physical or mental impairment, illness or 
disability (11%) or a special educational need (18%) compared with children in couple 
families (7% and 11% respectively).
The Children in Need (CIN) census is an annual survey from 2008-09 onwards that 
includes the numbers of disabled CIN in England. The CIN census enables DfE to 
support future policy development by achieving a better understanding of these 
vulnerable children. Latest figures show that of those recorded as children in need as at 
31 March 2022, 12.8% had a recorded disability. This figure has remained stable over 
the last few years.

### Housing status for adults with learning disabilities receiving 

support from social services
In the 2022 to 2023 financial year 80.5% of adults with learning disabilities aged 18-64 
who were receiving support from social services lived in their own home or with their 
families.
