---
output:
  govdown::govdown_document:
    font: "sans-serif" # (default) or "new-transport"
    favicon: "govuk" # or "none" (default) or "govuk" see ?govdown_document
    # logo: "images/logo.svg" # or false (default)
    logo_url: "https://ukgovdatascience.github.io/govdown/"
    logo_text: "GOV.UK"
    page_title: "Explore education statistics - SEN Key Statistics"
    title: "Explore education statistics"
    phase: beta
    feedback_url: "https://github.com/ukgovdatascience/govdown/issues"
    theme:
      bslib: true
---

``` {r include = FALSE}
library(rmarkdown)
library(bslib)
library(govdown)
knitr::opts_chunk$set(echo = FALSE)
```

## Headline statistics

``` {r}
bslib::layout_columns(
  col_widths = c(6,6),
bslib::value_box(
"Number of pupils with SEN requriements", 
  "1,766,924", 
theme="blue"
  ),
bslib::value_box(
"Percentage of pupils with SEN requirements", 
  "19.5%", 
theme="blue"
  )
)
```

Source: [Special educational needs in England, Academic year 2025](https://explore-education-statistics.service.gov.uk/find-statistics/special-educational-needs-in-england/2024-25)

## Overall SEN totals

``` {r}
bslib::layout_columns(
  col_widths = c(4, 4, 4),
  bslib::value_box(
    "Number of pupils on SEN support",
    "1,284,284",
    theme = "blue"
  ),
  bslib::value_box(
    "Number of Education, health and care plans at January 2025",
    "482,640",
    theme = "blue"
  ),
  bslib::value_box(
    "Total pupils with special education needs",
    "1,766,924",
    theme = "blue"
  )
)
```

Source: [Special educational needs in England, Academic year 2025](https://explore-education-statistics.service.gov.uk/find-statistics/special-educational-needs-in-england/2024-25)

## Type of need

Source: [Special educational needs in England, Academic year 2025](https://explore-education-statistics.service.gov.uk/find-statistics/special-educational-needs-in-england/2024-25)

## Characteristics

## Educational settings

## Absence and exclusions

## Outcomes and attainment

## Post-16 learner participation

## Destinations

## Experience of the SEND system
